Survival Compass: Statistical Insights into Lung Cancer Patients' Journey Post-Diagnosis

### 2024-04-04 paper/paper.qmd
finished making figures for data section. messing around to make them cuter :)

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

>>>>

```{r}
#| label: fig-bills
#| fig-cap: bb
#| echo: false

# install.packages
# install.packages("ggplot2")
# install.packages("dplyr")
# install.packages("showtext")

# load packaged
library(ggplot2)
library(dplyr)
library(showtext)

# 1. Add the desired font (e.g., Patrick Hand)
font_add_google("Patrick Hand", "patrick")

# 2. Set the font for ggplot
showtext_auto()

# Histogram of survival time in days (days_to_death)
ggplot(analysis_data, aes(x = days_to_death)) +
  geom_histogram(binwidth = 50, fill = "forestgreen", color = "black") +
  labs(title = "Distribution of Survival Time in Lung Cancer Patients Post-Diagnosis",
       x = "Survival Time (days)", y = "Frequency") +
  theme_minimal() +
  theme(
    text = element_text(family = "patrick") # Set font to Patrick Hand
  )

# create bar chart for pathogenic_stage
pathogenic_stage_counts <- analysis_data %>%
  group_by(pathogenic_stage) %>%
  summarise(count = n ()) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  arrange(desc(pathogenic_stage))

ggplot(pathogenic_stage_counts, aes(x = pathogenic_stage, y = percentage)) +
  geom_bar(stat = "identity", fill = "lightpink") +
  labs(title = "Percentage of Patients by Pathogenic Stage",
       x = "Pathogenic Stage",
       y = "Percentage") +
  theme_minimal() +
  theme(
    text = element_text(family = "patrick") # Set font to Patrick Hand
  )

###########################

# Define color palette for different stages
stage_colors <- c("Early" = "#FFB6C1", "Intermediate" = "#FF69B4", "Advanced" = "#FF1493")

# Modify data to ensure correct ordering of stages
analysis_data$pathogenic_stage <- factor(analysis_data$pathogenic_stage, levels = c("Early", "Intermediate", "Advanced"))

# Create bar chart for pathogenic_stage
pathogenic_stage_counts <- analysis_data %>%
  group_by(pathogenic_stage) %>%
  summarise(count = n ()) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  arrange(desc(pathogenic_stage))

# Plot the data
ggplot(pathogenic_stage_counts, aes(x = pathogenic_stage, y = percentage, fill = pathogenic_stage)) +
  geom_bar(stat = "identity") +
  labs(title = "Percentage of Patients by Pathogenic Stage",
       x = "Pathogenic Stage",
       y = "Percentage") +
  scale_fill_manual(values = stage_colors) +  # Apply color palette
  theme_minimal() +
  theme(
    text = element_text(family = "patrick")  # Set font to Patrick Hand
  )

# create bar chart for presence_of_distant metastasis 
presence_of_distant_metastasis_counts <- analysis_data %>%
  group_by(presence_of_distant_metastasis) %>%
  summarise(count = n ()) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  arrange(desc(presence_of_distant_metastasis))

ggplot(presence_of_distant_metastasis_counts, aes(x = presence_of_distant_metastasis, y = percentage)) +
  geom_bar(stat = "identity", fill = "gold") +
  labs(title = "Percentage of Patients by Presence of Distant Metastasis",
       x = "Presence of Distant Metastasis",
       y = "Percentage") +
  theme_minimal() +
  theme(
    text = element_text(family = "patrick") # Set font to Patrick Hand
  )

# create bar chart for lymph node involvement 
lymph_node_involvement_counts <- analysis_data %>%
  group_by(lymph_node_involvement) %>%
  summarise(count = n ()) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  arrange(desc(lymph_node_involvement))

ggplot(lymph_node_involvement_counts, aes(x = lymph_node_involvement, y = percentage)) +
  geom_bar(stat = "identity", fill = "cyan") +
  labs(title = "Percentage of Patients by Lymph Node Involvement",
       x = "Lymph Node Involvement",
       y = "Percentage") +
  theme_minimal() +
  theme(
    text = element_text(family = "patrick") # Set font to Patrick Hand
  )

# create bar chart for tumor size
tumor_size_counts <- analysis_data %>%
  group_by(tumor_size) %>%
  summarise(count = n ()) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  arrange(desc(tumor_size))

ggplot(tumor_size_counts, aes(x = tumor_size, y = percentage)) +
  geom_bar(stat = "identity", fill = "purple") +
  labs(title = "Percentage of Patients by Tumor Size",
       x = "Tumor Size",
       y = "Percentage") +
  theme_minimal() +
  theme(
    text = element_text(family = "patrick") # Set font to Patrick Hand
  )

# create bar chart for treatment decision
ggplot(treatment_decision_counts, aes(x = reorder(treatment_decision, -percentage), y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), hjust = -0.2, color = "black", size = 3) +
  labs(title = "Percentage of Patients by Treatment Decision",
       x = "Treatment Decision",
       y = "Percentage") +
  theme_minimal() +
  coord_flip() +
  theme(
    text = element_text(family = "patrick") # Set font to Patrick Hand
  )

# Create bar chart for treatment type
ggplot(treatment_type_counts, aes(x = reorder(treatment_type, -percentage), y = percentage)) +
  geom_bar(stat = "identity", fill = "orange") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), hjust = -0.2, color = "black", size = 3) +
  geom_text(data = treatment_type_counts, aes(label = treatment_type), hjust = 1.2, color = "black", size = 3) +
  labs(title = "Percentage of Patients by Treatment Type",
       x = "Treatment Type",
       y = "Percentage") +
  theme_minimal() +
  coord_flip() +
  scale_x_discrete(breaks = seq(0, max(treatment_type_counts$percentage), by = 5)) +
  theme(
    text = element_text(family = "patrick") # Set font to Patrick Hand
  )
